---
title: "Week 5 Problem Set"
author: "Lindsey Devir (https://github.com/ldevir9)"
date: "5/3/26"
output: pdf_document
---


1. Start a new document, problemset5.Rmd, in your soc10problemsets repository. (You can download a template [here](https://github.com/Soc10Spring2026/soc10problemsets_templates), but be sure to save it in your own soc10problemsets repo). Read the North Carolina shapefile into R using the process described in section \@ref(using-simple-features-in-r). Then plot the data using `ggplot2`. Use a variable other than `AREA` to fill the shapes, and try using one of the `scale_fill_` arguments to create a red color scale. See [these notes on color scales with ggplot](https://ggplot2-book.org/scales-colour.html) for some theory and ideas here. __Commit__ your changes. 

[Note: the scoring guidelines in brackets can be deleted before submitting. Load the nc data (0.5 points), use the correct function (0.5) and plot using ggplot2 with a variable other than AREA (1 point). Use scale_fill to create a red color scale (0.5 points).]

```{r q1, echo = TRUE}

library(sf)
library(ggplot2)

#load the data in
nc <- st_read(system.file("shape/nc.shp", package= "sf"))

#then plot using a variable (I used SID79)
ggplot(nc) +
  geom_sf(aes(fill = SID79)) +
  scale_fill_gradient(low = "salmon", high = "darkred")
  theme_minimal()

```

2. Use the `tidycensus` function to download income data for block groups in Santa Clara County. Plot the demographic data using `ggplot()` and comment on what it shows. __Commit__ your changes. 


[Download data for a correct variable in the correct geographies (1 point). Plot the map with ggplot (0.5 points). Describe findings accurately (0.5 points).]

This map illustrates median household income within block groups in Santa Clara County. For example, the high income areas are concentrated in specific areas of the county. In contrast, the lower income areas are in other specific regions of the county. Therefore, this hsows spatial inequality within the county because income levels are not evenly distributed with the area, but conversely show geopgraphic clustering. 

library(tidycensus)
library(ggplot2)

santa_clara_income <- get_acs(
  state = "CA", 
  county = "Santa Clara",
  geography = "block group", 
  variables = "B19013_001",
  geometry = TRUE,
  year = 2020
)

ggplot(santa_clara_income, aes(fill = estimate))+
  geom_sf(color = NA) +
  scale_fill_viridis_c(option = "magma") +
  theme_minimal()


3. Use the `tigris` function to download spatial data for counties in California. Plot the data using `ggplot()`. __Commit__ your changes. 

[Use tigris to gather boundaries at the correct geometries (0.5 points). Plot the data correctly (0.5 points).]

```{r q3, echo = TRUE, message = FALSE}

library(tigris)
library(ggplot2)

#get counties
CA_counties <- counties(state = "CA")

#plot
ggplot(CA_counties) +
  geom_sf() + 
  theme_minimal()

```

4. Download the clean AQI monitor datasets from Canvas ("aqi22_clean_ca.csv") or using the code below, and plot these inside the boundaries that you have downloaded from `tigris`. Then, calculate the average PM2.5 levels in each of the geographical units, and plot your geographies again according to these levels (as we did in section \@ref(spatially-joining-data)). Which areas had the best (and worst) air quality in 2022? What are the implications? __Commit__ your changes. 

[Create a plot with boundaries and AQI locations (0.5 points). Spatially join data (1 point) and calculate average PM2.5 levels in each unit (0.5 points). Plot the data again (0.5 points). Describe the findings and potential social implications (1 point).]

After using the data set it is clear that the worst air quality in within the Central Valley area/more southern part of California. Menaing, the areas with better air quality seem to be the ones along the coast/ more northern. Hence, this shows that pollution is not evenly distributed across areas. The implications of this is that those in these areas of worst air quality might be put at risk for an increased possibly of experiencing health risks, versus those in good air quality, which demonstrates unequal envioromental equality. 

```{r q4, echo = TRUE, message = FALSE}

library(readr)
library(RCurl)
library(sf)
library(dplyr)
library(ggplot2)
library(magrittr)
library(viridis)

# input url from github
url <- getURL("https://raw.githubusercontent.com/tylermcdaniel/css/main/Data/aqi22_clean_ca.csv")

# download aqi monitor data (from github)
aqi22m <- read_csv(url)

CA_counties <- counties(state = "CA")

aqi22 %<>%
  st_set_crs(st_crs(CA_counties))

aqi22 %<>%
  filter(`Pollutant Standard` == "PM25 24-hour 2012" & `State Name` == "California")

ggplot(aqi22) +
  geom_sf(data = CA_counties) +
  geom_sf(aes(col = `Arithmetic Mean`)) +
  scale_color_viridis(option = "magma", direction = -1) +
  theme minimal() +
  labs(col = "PM2.5 Average")

aqi_join <- st_join(CA_counties, aqi22)

aqi_counties <- aqi_join %<>%
  group_by(NAME) %<>%
  summarise(mn_aqi = mean(`Arithmetic Mean`, na.rm= TRUE))

ggplot(aqi_counties) +
  geom_sf(aes(fill = mn_aqi)) +
  scale_fill_viridis(option = "magma", direction = -1) +
  theme_minimal()
```


5. Does the county-level air quality map accurately represent how people experience air quality? Is the Modifiable Areal Unit Problem something that we should be concerned with? __Commit__ your changes. 

[Make a defensible argument about the data's representativeness (0.5 points). Make a defensible argument about the MAUP (1 point). ]

The county level map does not equally represent how people experience air quality because within the county itself the level of pollution can vary. Menaing, certain neighborhoods as we looked at in class might have more exposure than other areas. This can be that the monitors may be concentrated in one area, leaving other neighborhoods not measured to the same degree for its level of pollution. So when data is aggregated at the county level it is difficult to see and identify these outcome of spatial analysis can vary on the geographic unit used (such as county or neighborhood). In the exmaple we are talking about counties masked the data and made it hard to draw real conclusions by oversimplifying pattern, creating data unequal to all areas. This is very important because air pollution is correlated with health risks, so in using large spatial units like counties led to inequalities that can affect outcomes. Overall, MAUP can be somehting of concern as it is possible to lead to wrongful or simplified conclusions about envioromental issues.

