---
title: "Week 4 Problem Set"
author: "Lindsey Devir (https://github.com/ldevir9)"
date: "4/26/26"
output: pdf_document
---

1. Start a new document, problemset4.Rmd, in your soc10problemsets repository. (You can download a template [here](https://github.com/Soc10Spring2026/soc10problemsets_templates), but be sure to save it in your own soc10problemsets repo). In the chunk below, show how you would gather data on California's ethnoracial demographics for 2020 using `tidycensus`. (Note: in the chunk below, `eval` is set to false, so you won't see any results. That is ok! Just show the code. And remove your personal key). __Commit__ your changes.

[Text answer for question 1 here]

```{r q1, echo = TRUE, eval = FALSE}


```

2. In Bit by Bit, Salganik describes several possibilities and challenges of survey data in the digital age. Do you think any of these have changed since 2017 (when the book was written)? Pick a new type of survey or data collection, and describe its current possibilities and limitations. __Commit__ your changes. 

One challenge in survey data is the rise in nonresponse rates, which I think has definitely increased more way more from 2017. The outcome threatens the quality of estimates because the estimates highly depend on the statistical models that researchers use to adjust for missing nonresponses. The outcome is that even probability sampling may not generate reliable results. Furthermore, a challenge in digital data specifically like non-probability sampling is unadjusted samples can lead to bad estimates, they are cheaper and faster to conduct but are less familiar and more varied resulting in unrepresentative samples. With phones becoming prominent I think it could be intresting to study people's social media data to track behaviors in what one interacts with. I think a challenge of this would be people feeling like their information is being exposed, and used without consent, causing issues reguaring privacy when using public data. Furthermore, this might not be fully a representative sample since part of the population might not have social media, or even a phone. However, the bonus of this sample method is that it allows to gather large amounts of data to draw information of how people interact with information, studying social behavior. Hence, more digital data collection methods could create new forms of research but also create new possibilities for bias or limitations for unrepresentative samples.


3. Gather the data from [this Wikipedia page on incarceration in the US](https://en.wikipedia.org/wiki/Incarceration_in_the_United_States) (you can use the code from Section 4 of the course website to gather this). Use `mutate()`, `parse_number()` from the `readr` package, and the magrittr pipe (`%<>%`) to fix the issue with the `Count` variable. Then plot the Count variable over time with `ggplot()`. __Commit__ your changes.

[Text answer for question 3 here]

```{r q3, echo = TRUE, message = FALSE}



```

4. Now model it! Add a linear model of incarceration over time to your plot. (Note: If you are unable to plot question 3 correctly, you can plot `Rate` as the outcome for this question). __Commit__ your changes.  

[Text answer for question 4 here]

```{r q4, echo = TRUE}


```


5. Would a linear or non-linear model be more appropriate for explaining and predicting trends in US society and incarceration? Explain the benefits and drawbacks of each. __Commit__ your changes.  

[Text answer for question 5 here]

